---
title: "ojoThemes Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ojothemes-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This guide will walk you through the themes and tools available in the `ojoThemes` package. We'll start by loading the package (plus `tidyverse` and `ojodb` for some test data), then we'll make some example graphs and tables to show you what's available.

```{r setup}
library(ojothemes)
library(ojodb)
library(tidyverse)
library(gt)

# Test data from OCDC
data <- ojo_tbl(schema = "ocdc",
                table = "arrest") |>
  filter(book_date >= "01-01-2024",
         book_date < "11-01-2024") |>
  ojo_collect() 

```
## Themes for `ggplot` graphs

First, let's see how this data looks on a default ggplot:

```{r default_ggplot}
p1 <- data |>
  # We'll look at the bookings by race here,
  # and we'll look by month just to make it easier to see.
  count(book_month = floor_date(book_date, "months"), 
        race) |>
  ggplot(aes(x = book_month, y = n, fill = race)) +
  geom_col()

p1
```

Next, let's gussy this up with our `ojoThemes` tools. First, let's use the `ojothemes::ojo_labs()` function to easily add a nice caption to the plot:

```{r ojo_labs}
p1 <- p1 +
  ojo_labs(analyst_name = "Andrew Bell",
           source = "ocdc",
           title = "OCDC Bookings by Race",
           subtitle = "January 2024 - October 2024",
           x = "Month",
           y = "Total Booking Events")

p1
```

This is just a wrapper for the normal `ggplot2::labs()` function, so we can pass arguments like `title`, `subtitle`, `x`, and `y` to it. The really neat thing, though, is the new `analyst_name` and `source` arguments. These will add a nice little "Graphic created by..." caption crediting the analyst, and a consistently phrased note on where the data came from.

  * Right now, you can use `"oscn"`, `"ocdc"`, or `"ppb"` as the `source` argument to get a pre-written, consistent source note. However, if you're working with something not in one of those domains, you can just tell it something to put in there verbatim, e.g. `source = "Source: Data Pulled from The National Insitute of My Ass database"`.

Next, let's actually apply our themes. There are two ways to do this: the first is to just throw `+ theme_okpi()` onto our ggplot, like so:

```{r plus_theme_okpi}
p1 + theme_okpi()
```

This applies not just the theme / fonts, but the color / fill scales as well. If you want just the theme without the scales, you can do `p1 + theme_okpi_base()`.

The second way of applying a theme is by using `ojothemes::ojo_set_theme()`. This is probably going to be the most common way of doing things, since we're rarely going to want to use multiple themes in the same script / document / etc. Let's try that same graph in the "ojo" theme this time:

```{r set_theme_okpi}
ojo_set_theme(theme = "ojo")

p1
```

## Themes for `gt` tables

This package also has built in themes for tables made with `gt`. You can use these by replacing the normal `gt()` function with one of our themed wrappers (currently `gt_okpi()` and `gt_ojo()`). These have built in arguments for customization: 

  * `source` and `analyst_name` (which work just like the `ojo_labs()` function for ggplots)
  * `title` and `subtitle` (self explanatory)
  * `format_cols` (a logical that will apply the auto-styling from `gt` to every column if `TRUE`)
  * `font` and `font_size` for making really big / small tables.

Here's what a normal, unstyled table looks like...

```{r gt}
data |>
  count(race) |>
  gt()
```

...here's what our `gt_okpi()` theme does to it...

```{r gt_okpi}
data |>
  count(race) |>
  gt_okpi(title = "OCDC Bookings by Race",
          subtitle = "January 2024 - October 2024",
          source = "ocdc",
          analyst_name = "Andrew Bell",
          format_cols = TRUE)
```

...and finally, here's `gt_ojo()`:

```{r gt_ojo}
data |>
  count(race) |>
  gt_ojo(title = "OCDC Bookings by Race",
         subtitle = "January 2024 - October 2024",
         source = "ocdc",
         analyst_name = "Andrew Bell",
         format_cols = TRUE)
```
